DATA 621 Homework #4

### 2019-11-05 hw4/HW4.Rmd
Updated HW4

---
title: "DATA 621 Homework #4"
author: "Critical Thinking Group 3"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(dplyr)
library(Amelia)
library(kableExtra)
library(caret)
library(DMwR)
# Thank you Stack Overflow!
# A Prefix nulling hook.

@@

```{r}
df <- read.csv('./data/insurance_training_data.csv') %>%
  mutate(TARGET_FLAG = as.factor(TARGET_FLAG))
evaluation <- read.csv("./data/insurance-evaluation-data.csv")
```

@@

### Are There Missing Values?
First we look at the data to see if any variables have missing data:

@@

There are missing values for the `CAR_AGE` and `YOJ`(Years on Job) variables.  We will fill in the missing variables using KNN.

```{r}
set.seed(42)
knn <- knnImputation(df)
impute_me <- is.na(df$CAR_AGE)
df[impute_me,"CAR_AGE"] <- knn[impute_me,"CAR_AGE"] 
impute_me <- is.na(df$YOJ)
df[impute_me,"YOJ"] <- knn[impute_me,"YOJ"] 
```

### Splitting the Data
Next, we look to split our data between a training set (`train`) and a test data set (`test`). We'll use a 70-30 split between train and test, respectively.

@@

### First Look at the Data

We will now take our first look at the data to get a sense of what we have.

```{r}
train %>%
  summary() %>%
  kable() %>%
  kable_styling()
```

>>>>

---
title: "DATA 621 Homework #4"
author: "Critical Thinking Group 3"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(Amelia)
library(kableExtra)
library(caret)
library(DMwR)
# Thank you Stack Overflow!
# A Prefix nulling hook.

@@

```{r}
df <- read.csv("./data/insurance_training_data.csv") %>%
  mutate(TARGET_FLAG = as.factor(TARGET_FLAG)) %>%
  select(-INDEX)
evaluation <- read.csv("./data/insurance-evaluation-data.csv") %>%
  select(-INDEX)
```

## Introduction

We have been given a dataset with `r nrow(df)` records representing customers of an auto insurance company.  Each record has two response variables.  The first is a binary flag where one means a person was in a car crash and zero means they were not.  There are `r df %>% filter(TARGET_FLAG == 1) %>% nrow()` rows with a one flag and `r df %>% filter(TARGET_FLAG == 0) %>% nrow()` with a zero flag.  
The first objective is to logistic regression classifier to predict if a person was in a car crash.  The second sreponse variable is the amount it will cost if the person crashes their car.  The value is zero if the person did not crash their car.

@@

## Data Preparation

```{r}
df %>%
  summary() %>%
  kable() %>%
  kable_styling()
```

### Fix Data Types

There are some variables that are currently factors that are dollar values that need to be transformed into numeric variables.  We will do this to both the `df` and `evaluation` data frames.

```{r}
strip_dollars <- function(x){
  x <- as.character(x)
  x <- gsub(",", "", x)
  x <- gsub("\\$", "", x)
  as.numeric(x)
}

fix_data_types <- function(messy_df){
  messy_df %>%
    rowwise() %>%
    mutate(INCOME = strip_dollars(INCOME),
           HOME_VAL = strip_dollars(HOME_VAL),
           BLUEBOOK = strip_dollars(BLUEBOOK),
           OLDCLAIM = strip_dollars(OLDCLAIM)) %>%
    ungroup()
}

df <- fix_data_types(df)
evaluation <- fix_data_types(evaluation)
```

@@

```{r}
df %>%
  summary() %>%
  kable() %>%
  kable_styling()
```

### Fix Missing Values
There are missing values.  We will fill in the missing variables using KNN.

```{r, null_prefix = TRUE}
set.seed(42)
#knn <- df %>%
#  select(-TARGET_FLAG, -TARGET_AMT) %>%
#  na.omit() %>%
#  knnImputation()

for(var in names(df)){
  impute_me <- is.na(df[[var]])
  if (nrow(df[impute_me,]) > 0){
    print(paste("Fixing the", nrow(df[impute_me,]), "missing values in", var, "(df)"))
    #df[impute_me, var] <- knn[impute_me, var] 
  }
}
```

### Feature Creation

We will create a log transformed income and home value feature.  We will also create an average claim amount that will hopefully track better with the `TARGET_AMT` variable.

```{r}
feature_creation <- function(d){
  d %>%
    rowwise() %>%
    mutate(LOG_INCOME = log(INCOME + 1),
           LOG_HOME_VAL = log(HOME_VAL + 1),
           AVG_CLAIM = ifelse(CLM_FREQ > 0, OLDCLAIM / CLM_FREQ, 0)) %>%
    ungroup()
}

df <- feature_creation(df)
evaluation <- feature_creation(evaluation)
```

### Creating Training/Test Data Sets

Now that we have a complete data set we will split the data into a training (`train`) and test set (`test`). We'll use a 70-30 split between train and test, respectively.

@@
